% Appendix A

\chapter*{Acknowledgements} % Main appendix title

The writing of this manuscript closes a period of almost four years dedicated to science. It has made me grow both 
personally and technically and I am very grateful for the opportunity. In the way, I have had the chance to meet 
formidable scientists, technicians, and friends, which I will try to mention in the following paragraphs. \\

I started tasting science in my bachelor thesis with Aitor and Ion. I continued working with them in the Master 
thesis and we have worked together in the Ph.D. thesis. I think we have had the chance to make fascinating science. I 
could not have chosen better supervisors. You are not only passionate scientists, but also honest people who I 
appreciate. Obviously, this work would not have been possible without your support and advice. I would specially like 
to thank Ion for his daily and close dedication. \\

I would like to acknowledge my current institute Centro de F\'isica de Materiales and all the people working in it. I 
would also like to thank the Donostia International Physics Center (DIPC) and its computing center service, as part 
of the computational work presented in this thesis was done on the DIPC supercomputers. The rest of the computational
resources where provided by Marenostrum and PRACE. The financial support was provided by the Material Physics 
Center. \\

In the second year of my Ph.D. I had the opportunity to spend three months working with Matteo Calandra and Lorenzo 
Paulatto in Paris. I appreciate their warm welcome and all the advices and teaching they gave me in those three 
months. I think it was a very fruitful period of the thesis were I learned many new skills. \\

I have had the privilege to share office and building with wonderful colleagues. I need to start with Miguel, who 
really dedicated time and effort on resolving my technical and conceptual issues at the beginning of my scientific 
travel. Apart from that we shared many coffee breaks and paddle matches together with Toma\^s, Alba, Jose, Joseba, 
Julen, Mikel, Raffaello, Antonnela, Francesco, Martin, Antton, and many others who I probably forget. It has been a 
pleasure to spend this time with all of you. \\

Finally I want to thank my family and friends for the support during all these years. You have always shown interest 
on how I was doing, which has given me strength to continue doing my job. It has been important for me to see that 
close people believed on what I was doing. I want to finish with Olaia. You have been my partner during all these 
years and we have built this work together. There has been no moment where I have not felt your support and love. 
Even today you keep asking how graphene or SnSe are going. All these years would have been completely different 
without your company, eskerrik asko. 
